---
title: "R5-HypothesisTesting"
author: "OGabriel"
date: "03/11/2020"
output: 
  html_document:
     toc: yes
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
  fig.path = "results/figs/",       # Path to save figures
  dev = "png",                      # png Figures format
  dpi = 100  
)

#install.packages("agricolae")
```

## Exercise 1: 
For the data InsectSpray, make a table for the number of insects for each spray with the mean, median and standard error.

###  Summary of the data InsectSpray for a general overview of the data

```{r, echo=FALSE}
data("InsectSprays")
summary(InsectSprays)

Spray <- character()
Mean <- Standard_error <- Median <-Insect_Number <-  numeric()
for (x in levels(InsectSprays$spray)){
    Spray <- c(Spray, x)
    j <- InsectSprays$count[InsectSprays$spray == x]
    Insect_Number <- c(Insect_Number, sum(j))
    Mean <- round(c(Mean, mean(j)), 2)
    Standard_error <- round(c(Standard_error,sd(j) / sqrt(length(j))), 2)
    Median <- c(Median, median(j))
}
###round(,2) was used to reduce the mean and standard error to 2 decimal points
knitr::kable(D <- data.frame(Spray, Insect_Number,  Mean, Standard_error, Median), caption = "Table 1. Insect count as a result of spray")
```






## Exercise 2: 
Print a plot to see the differences of counts between sprays. Include a caption explaining the figure. Which type of plot is the one you choose and why?

```{r,fig.cap= "Figure 1. Differences of counts between sprays",echo=FALSE}

data("InsectSprays")
plot(count ~ spray, data = InsectSprays, 
     main = "Differences of count between sprays")

##Boxplot was chosen because it shows how well distributed the data is dividing the data set into three quartiles (the minimum, maximum, median, first quartile 
```






## Exercise 3: 
Test for differences between sprays using anova and a posthoc comparison and redo the previous plot including the representation of all posthoc differences.

Note: for the anova use the command aov() and for the posthoc comparison use the Tukey’s ‘Honest Significant Difference’ method. For this method try the TukeyHSD() and the agricolae::HSD.test() and see the differences.


###  ANOVA: res.aov1
```{r, echo=FALSE}
res.aov1 <- aov(count ~ spray, data = InsectSprays)
summary(res.aov1)

```


###  TukeyHSD
```{r, echo=FALSE}
tukey.test1 <- TukeyHSD(res.aov1)
print(tukey.test1)
```

###  agricolae::HSD.test
```{r, fig.cap= "Figure 2. HSD.test for res.aov1", echo=FALSE}
library(agricolae)
ASD1 <- HSD.test(res.aov1, "spray", console = TRUE)
##console = TRUE to display the result in the console

###plot including the representation of all posthoc differences using HSD.test method
plot(count ~ spray, data = InsectSprays, ylim = c(0,30), 
     main = "HSD.test for res.aov1")
text(x = 1:6 , y = 29, c("a", "a", "b", "b", "b", "a"))

```



## Exercise 4: 
Test for differences between sprays using non-parametric 
Kruskal-Wallis rank sum test. Again, redo the plot with these results. Use agricolae::kruskal()

###  Summary
```{r, fig.cap= "Figure 3. Kruskal-Wallis rank sum test between sprays", echo=FALSE}
library(agricolae)
data("InsectSprays")
str(InsectSprays)
krus <- kruskal(InsectSprays$count, trt = InsectSprays$spray)
print(krus)
 
##Making a plot for the easy presentation of the kruskal comparison
plot(count ~ spray, data = InsectSprays, ylim = c(0,30), 
     main = "Kruskal-Wallis rank sum test between sprays")
text(x = 1:6 , y = 29, c("a", "a", "c", "b", "bc", "a"))

```
















##  Exercise 5:
Transform count data using sqrt(count) and redo the anova, the Tukey posthoc comparison and the plot.

###  Transforming count data of InsectSprays using sqrt(count)
```{r, echo=FALSE}
sqct <- sqrt(InsectSprays$count) 
sqct
```

###  ANOVA after count transformation using sqrt(count): res.aov2
```{r, echo=FALSE}

res.aov2 <- aov(sqct ~ spray, data = InsectSprays)
summary(res.aov2)

```

### TukeyHSD
```{r, echo=FALSE}
tukey.test2 <- TukeyHSD(res.aov2)
print(tukey.test2)

```

###  agricolae::HSD
```{r,fig.cap=  "Figure 4.HSD:test for res.aov2", echo=FALSE}
ASD2 <- HSD.test(res.aov2, "spray", console = TRUE)
##console = TRUE to display the result in the console


##plot of HSD:test after sqrt transformation
plot(count ~ spray, data = InsectSprays, ylim = c(0,30), 
     main= "HSD:test for res.aov2")
text(x = 1:6 , y = 29, c("a", "a", "c", "b", "bc", "a"))



```



## Exercise 6:
Test for normality of residuals for the two performed anova analyses of points 4 and 6 using shapiro.test() and use plot the anova to see the qqplots and compare them.

###  Shapiro.test for the ANOVA res.aov1 
```{r, echo=FALSE}
shapiro.test(res.aov1$residuals)

```

###  Shapiro.test for the ANOVA res.aov2
```{r,echo=FALSE }
shapiro.test(res.aov2$residuals)
```

###  QQplot for comparison
```{r,fig.cap= "Figure 5. QQplot for comparison for res.aov1 and res.aov2", echo=FALSE}
qqnorm(res.aov1$residuals, main = "QQplot for res.aov1")
qqline(res.aov1$residuals)
qqnorm(res.aov2$residuals, main = "QQplot for res.aov2")
qqline(res.aov2$residuals)

```

## Exercise 7:
###  Which of the previous analysis is the adequate in this case? Why? 

The ANOVA with the transformed count is adequate.It allowed us to prove normality. It was more accurate than the direct ANOVA into detecting significant differences between sprays. 
For example, the diffence between spray C and spray D was detected. 

###  Is there any difference in the results between the square root transformed ANOVA and the Kruskal-Wallis analyses? 
There was no difference in the transformed ANOVA and the Kruskal-Wallis analyses. Both presented the same significance difference between the sprays using the alphabets. 

###  Is there any difference in the results between the direct ANOVA and the square root transformed ANOVA? Which ones?
There is a difference between the two ANOVAs. The square root transformed ANOVA (res.aov2) exposed the statistical difference between spray C and spray D which was not shown in the direct ANOVA (res.aov1)









